---
output: github_document
---


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```




## Intro {ggedgelist} idea

Here we think about a shortcut to a first looks at networks using a flat, edgelist input that leads straight to a ggplot2 plot (ggraph) space; this uses {tidygraph} and {ggraph} under the hood...



## Step 00 Before getting into it, create an 'interesting' edge list

```{r cars, message=F, warning=F}
library(tidyverse)
library(tidygraph)
library(ggraph)

set.seed(12345)

edge_list <- 
  data.frame(node_to = sample(rep(LETTERS[1:10], 50), 
                              replace = T), 
             node_from = sample(rep(LETTERS[1:10], 50),
                                replace = T)) %>% 
  arrange(node_to, node_from) %>% 
  count(node_to, node_from) %>% 
  sample_n(18)

head(edge_list)
```

## Step 0. Whats the status quo edgelist -> network viz

```{r}
edge_list %>% 
  as_tbl_graph() %>% 
  ggraph() +
  geom_edge_link(color = "orange") +
  geom_node_point(size = 9,
                  color = "steelblue",
                  alpha = .8) + 
  geom_node_text(aes(label = name))

# aesthetic mapping for edge characteristics
last_plot() + 
  geom_edge_link(color = "red", 
                 alpha = .5 ,
                 aes(edge_width = n))
```

## Step 0.a And to also vizualize node atributes


```{r}
node_info <- data.frame(my_nodes = LETTERS[1:10], 
                        ind_child = sample(c(T,F), 10, 
                                           replace = T))

edge_list %>% 
  as_tbl_graph() %>%
  left_join(node_info %>% 
              rename(name = my_nodes)) %>%
  ggraph() +
  geom_edge_link(color = "orange") +
  geom_node_point(size = 9,
                  color = "steelblue",
                  alpha = .8) + 
  geom_node_text(aes(label = name))

```



# Proposed functions, edgelist -> plotspace feel


```{r ggedgelist}
# get into ggplot2 plot space from edge list data frame 
ggedgelist <- function(edgelist, nodelist = NULL, ...)(
  
  # message("'name' a variable created in the 'nodes' dataframe")
  
    if(is.null(nodelist)){
    edgelist %>% 
    tidygraph::as_tbl_graph() %>% 
    ggraph::ggraph(...) 
    
  }else{ # join on nodes attributes if they are available
    
    names(nodelist)[1] <- "name"
    
    edgelist %>% 
    tidygraph::as_tbl_graph() %>%
    dplyr::full_join(nodelist) %>% 
    ggraph::ggraph(...) 
    
  }
  
)

# get a fill viz w edgelist dataframe only
ggedgelist_quick <- function(edgelist, nodelist = NULL, include_names = F,  ...){
  

  p <- ggedgelist(edgelist = edgelist,
                  nodelist = nodelist, ...) +
  ggraph::geom_edge_link(color = "orange") +
  ggraph::geom_node_point(size = 9,
                  color = "steelblue",
                  alpha = .8) 
  
  if(include_names){p + ggraph::geom_node_label(aes(label = name))}else{p}
  
}

geom_node_label_auto <- function(...){ 
  
  ggraph::geom_node_label(aes(label = name), ...)
  
}

geom_node_text_auto <- function(...){ 
  
  ggraph::geom_node_text(aes(label = name), ...)
  
}
```





# examples w/ proposed functions


## `ggedgelist_quick()`

```{r, out.width="33%", fig.show='hold', message=F}
head(edge_list)
head(node_info)
edge_list %>% 
  ggedgelist_quick() 

edge_list %>% 
  ggedgelist_quick(include_names = T) 

edge_list %>% 
  ggedgelist_quick(nodelist = node_info) + 
  geom_node_point(aes(color = ind_child), size = 10)
```

## `ggedgelist()` + `geom_node_label_auto()`

```{r}
edge_list %>% 
  ggedgelist(layout = "kk") + 
  geom_edge_link(arrow = arrow(), linetype = "dashed") +
  geom_node_tile(width = .18, height = .5) + 
  geom_node_label_auto() 
```

# using the ggflowchart example (100% inspiration)

```{r}
# flowcharter example
ggflowchart_example <- tribble(~from, ~to,
        "A", "B",
        "A", "C",
        "A", "D",
        "B", "E",
        "C", "F",
        "F", "G") 

ggflowchart_example %>% 
  as_tbl_graph()
```


## Start by using quick plot function 'qedgelist2ggraph'

```{r, out.width="33%", fig.show='hold', message=F}
ggflowchart_example %>% 
  ggedgelist_quick(layout = "stress", 
                   include_names = T)

ggflowchart_example %>% 
  ggedgelist_quick(layout = "tree", 
                   include_names = T)

# auto which is default also produces three in this case
ggflowchart_example %>% 
  ggedgelist_quick(layout = "auto") 

layer_data(last_plot(), i = 2)
```

## Use ggedgelist and geom_edge_* and geom_node_* functions to customize

```{r}
ggflowchart_example %>% 
  ggedgelist() +
  geom_edge_link(linetype = "dashed") + 
  geom_node_point(size = 12, alpha = .2) + 
  geom_node_label_auto(fill = "magenta")

```

## Capabilities end here.  If you need to access powerful network calculation capabilities, step back into the tidygraph world!  


```{r}
ggflowchart_example %>% 
  as_tbl_graph() %>% 
  mutate(dg_cent = centrality_degree()) %>% 
  ggraph("stress") +
  geom_edge_link(linetype = "dashed") + 
  geom_node_point(alpha = .2, aes(size = dg_cent)) + 
  scale_size(range = c(8, 15)) +
  geom_node_label_auto(fill = "magenta")

```



# Part II. Packaging and documentation  🚧 ✅ 



## Phase 1. Minimal working package

### Created files for package archetecture with `devtools::create(".")` ✅ 

### Moved functions R folder? ✅  

```{r}
knitr::knit_code$get() |> names()
```

Use new {readme2pkg} function to do this from readme... ✅ 

```{r}
readme2pkg::chunk_to_r("ggedgelist")
```


### Added roxygen skeleton? ✅ 

Use a roxygen skeleton for auto documentation and making sure proposed functions are *exported*.

### Managed dependencies ? ✅ 

Package dependencies managed, i.e. `depend::function()` in proposed functions and declared in the DESCRIPTION

```{r pkg_dependencies}
usethis::use_package("ggplot2")
usethis::use_package("ggraph")
usethis::use_package("tidygraph")
usethis::use_package("dplyr")
```



### Chosen a license? ✅ 


```{r pkg_license}
usethis::use_mit_license()
```

### Run `devtools::check()` and addressed errors? 🚧 

```{r pkg_check, results='hide', error=T}
devtools::check(pkg = ".")
```



### Build package 🚧 

```{r pkg_build}
devtools::build()
```

You need to do this before library(mynewpackage) will work.

### Make aspirational part of readme real. 🚧 

At this point, you could change eval chunk options to TRUE. You can remove  the 🦄 emoji and perhaps replace it with construction site if you are still uncertain of the API, and want to highlight that it is subject to change. 

### Add lifecycle badge (experimental)✅  

```{r pkg_lifecycle_badge}
usethis::use_lifecycle_badge("experimental")
```


## Phase 2: Listen & iterate 🚧 

Try to get feedback from experts on API, implementation, default decisions.  Is there already work that solves this problem? 


## Phase 3: Let things settle

### Settled on examples.  Put them in the roxygen skeleton and readme. 🚧 

### Written formal tests of functions? 🚧 


That would look like this...

```{r test_calc_frequency_works, eval = F}
library(testthat)

test_that("calc frequency works", {
  expect_equal(calc_frequency("A", 0), 440)
  expect_equal(calc_frequency("A", -1), 220)
})
```


```{r send_tests, eval = F}
readme2pkg::chunk_to_tests_testthat("test_calc_frequency_works")
```


### Have you worked added a description and author information in the DESCRIPTION file? 🚧 

### Addressed *all* notes, warnings and errors. 🚧 

## Promote to wider audience...

### Package website built? 🚧 

### Package website deployed? 🚧 

## Phase 3: Harden/commit

### Submit to CRAN? Or don't.  🚧 

# Appendix: Reports, Environment

## Description file extract

```{r}

```

## Environment

Here I just want to print the packages and the versions

```{r session_pkgs}
all <- sessionInfo() |> print() |> capture.output()
all[11:17]
```

## `devtools::check()` report

```{r report_check, error = T, results="hide", warning=F}
devtools::check(pkg = ".")
```





